---
title: "p8105_hw6_sal2222"
author: "Stephen Lewandowski"
date: "November 16, 2018"
output:  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

### Problem 1
This problem utilizes data from _The Washington Post_ on homicides in 50 U.S. cities. The code chunk below imports and cleans the data. I referred to the posted solution of Homework 5, Problem 2 for some of the data cleaning code structure and naming conventions.

```{r}
homicide_df <- 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>%
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolution = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest"        ~ "unsolved",
      disposition == "Closed by arrest"      ~ "solved"
    ),
    victim_race = case_when(
      victim_race == "White" ~ "white",
      victim_race != "white" ~ "non-white"
    ),
    victim_age = as.numeric(victim_age)) %>% 
  filter(!city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO")) 
```



In the next code chunk, I group within cities and summarize to produce the total number of homicides and the number that are solved. 

```{r}
city_homicide_df = 
  homicide_df %>% 
  select(city_state, disposition, resolution) %>% 
  group_by(city_state) %>% 
  summarize(hom_total = n(),
            hom_unsolved = sum(resolution == "unsolved"))
```

Focusing only on Baltimore, MD, I can use the `prop.test` and `broom::tidy` functions to obtain an estimate and CI of the proportion of unsolved homicides in that city. The table below shows those values.

```{r}
bmore_test = 
  prop.test(x = filter(city_homicide_df, city_state == "Baltimore, MD") %>% pull(hom_unsolved),
            n = filter(city_homicide_df, city_state == "Baltimore, MD") %>% pull(hom_total)) 

broom::tidy(bmore_test) %>% 
  knitr::kable(digits = 3)
```


