---
title: "p8105_hw6_sal2222"
author: "Stephen Lewandowski"
date: "November 16, 2018"
output:  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(MASS)
library(modelr)
library(mgcv)
library(tidyverse)


theme_set(theme_bw() + theme(legend.position = "bottom"))
```

### Problem 1 - Homicide data

This problem utilizes data from _The Washington Post_ on homicides in 50 U.S. cities. The code chunk below imports and cleans the data. I referred to the posted solution of Homework 5, Problem 2 for some of the data cleaning code structure and naming conventions.

```{r clean_homicide, message = FALSE}
homicide_df <- 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>%
  filter(victim_race != "NA") %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolution = factor(case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest"        ~ "unsolved",
      disposition == "Closed by arrest"      ~ "solved"
    )),
    victim_race = factor(case_when(
      victim_race == "White" ~ "white",
      victim_race != "white" ~ "non-white"
    )),
    victim_race = fct_relevel(victim_race, "white"),
    victim_age = as.numeric(victim_age),
    victim_sex = factor(victim_sex)) %>% 
  filter(!city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO")) 
```

I excluded cities that do not report race as well as a data entry error ("Tulsa, AL"). I also excluded 4,199 observations with `victim_race` listed as `NA`. 

## Baltimore model

For the city of Baltimore, MD, I will apply the glm function to fit a logistic regression with `resolved` versus `unresolved` as the outcome and victim age, sex and race as predictors. The output is saved as R object `fit_logistic`. 

```{r glm_baltimore}

fit_logistic_baltimore <- 
  homicide_df %>% 
    filter(city_state == "Baltimore, MD") %>% 
    dplyr::select(resolution, victim_age, victim_race, victim_sex) %>% 
    glm(resolution ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) 

```

I will apply `broom::tidy` to `fit_logistic` and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims, keeping all other variables fixed.

```{r broom_baltimore}

broom::tidy(fit_logistic_baltimore, conf.int = TRUE) %>% 
  mutate(OR = exp(estimate),
         conf_low_OR = exp(conf.low),
         conf_high_OR = exp(conf.high)) %>%
  dplyr::select(term, OR, conf_low_OR, conf_high_OR) %>%
  filter(term == "victim_racenon-white") %>% 
  knitr::kable(digits = 3)


```


In Baltimore, homicides of non-white victims were 2.27 times more likely to be unsolved compared to homicides of white victims after adjusting for age and gender.

## All cities

Now, I will run a glm for each of the cities in my dataset and extract the adjusted odds ratio and CI for solving homicides comparing non-white victims to white victims within a “tidy” pipeline.


```{r glm_all_cities}

nest_glm_all <-
  homicide_df %>% 
  group_by(city_state) %>% 
  nest() %>% 
  mutate(models = map(data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = ., family = binomial())),
         models = map(models, broom::tidy, conf.int = TRUE)) %>% 
  select(-data) %>% 
  unnest()

```

From these resuts, I will create a dataframe with the estimated, adjusted ORs and CIs for each city comparing unsolved cases among non-white and white victims.

```{r ORs_all_cities}
all_city_OR <-
  nest_glm_all %>% 
    mutate(OR = exp(estimate),
           conf_low_OR = exp(conf.low),
           conf_high_OR = exp(conf.high)) %>%
    dplyr::select(city_state, term, OR, conf_low_OR, conf_high_OR) %>%
    filter(term == "victim_racenon-white")  

```

The results are displayed as a `kable` table below.
```{r}
all_city_OR %>%
   knitr::kable(digits = 3)
```

I will now plot the estimated ORs and CIs for each city.

```{r OR_plot}

all_city_OR %>%
  mutate(city_state = forcats::fct_reorder(city_state, OR)) %>%
  ggplot(aes(x = city_state, y = OR)) + 
      geom_errorbar(aes(ymin = conf_low_OR, ymax = conf_high_OR), width = 0.2) +
      geom_point(size = 3, shape = 21, fill = "white") + 
  labs(
    title = "Unsolved homicides, comparing non-white victims to white victims",
    x = "City",
    y = "Adjusted odds ratio",
    caption = "Data from the Washington Post, adjusted for victim age, race, and gender."
  ) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
```

The results and plot indicate that the U.S. cities with the largest discrepencies in unsolved homicide case by victim race, after adjusting for victim age and gender are Boston, Omaha, Oakland, Pittsburgh, and Cincinnati. Boston also has the largest confidence interval about the adjusted OR. 


### Problem 2 - Birthweight data

I will now load and clean birthweight data for 4,342 children for regression analysis.

```{r}
birthweight_df <- 
  read_csv("data/birthweight.csv") %>% 
  mutate(
      babysex = factor(babysex),
      frace = factor(frace),
      malform = factor(malform),
      mrace = factor(mrace)
  )
```

There is no missing data in the dataframe.

```{r birthweight_na_check}
birthweight_df %>% 
  is.na() %>% summary()
```

## Proposed regression model for birthweight: smoking and maternal age

I propose a model for birthweight based on a hypothesized structure. I am interested in seeing the effects of smoking as an environmental exposure as well as maternal age. I will examine the fit of `smoken` and `menarche` independently, jointly, and with interaction.


First, I will review the simple linear regression estimates and p-values.
```{r my_models_broom_tidy}
lm(bwt ~ smoken, data = birthweight_df) %>% broom::tidy()
lm(bwt ~ menarche, data = birthweight_df) %>% broom::tidy()
lm(bwt ~ smoken + menarche, data = birthweight_df) %>% broom::tidy()
lm(bwt ~ smoken + menarche + (smoken * menarche), data = birthweight_df) %>% broom::tidy()
```

Next, I will review model fit statistics.
```{r my_models_broom_glance}
lm(bwt ~ smoken, data = birthweight_df) %>% broom::glance()
lm(bwt ~ menarche, data = birthweight_df) %>% broom::glance()
lm(bwt ~ smoken + menarche, data = birthweight_df) %>% broom::glance()
lm(bwt ~ smoken + menarche + (smoken * menarche), data = birthweight_df) %>% broom::glance()
```

Based on considration of parameter estimate results and goodness of fit tests (including maximized adjusted R-squared and minimized AIC), I select the model with both average number of cigarettes smoked per day during pregnancy and mother’s age at menarche (years). There is not significant interaction between these terms.

My selected model is: `lm(bwt ~ smoken + menarche)`.
```{r}
smoken_menarche_fit <-
  lm(bwt ~ smoken + menarche, data = birthweight_df)
```

## Residuals plot

Below is a plot of model residuals against fitted values.

```{r residuls_plot}

smoken_menarche_resid <-
  modelr::add_residuals(birthweight_df, smoken_menarche_fit)

smoken_menarche_pred <-
  modelr::add_predictions(birthweight_df, smoken_menarche_fit)

smoken_menarche_full <-
  birthweight_df %>% 
    mutate(resid = smoken_menarche_resid$resid,
           pred = smoken_menarche_pred$pred)
  
smoken_menarche_full %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  geom_smooth() +
  labs(
    title = "Proposed model residuals against fitted values",
    x = "Predictions",
    y = "Residuals",
    caption = "Model: lm(bwt ~ smoken + menarche, data = birthweight_df)"
  ) 

```

In this plot, the range of residuals tends to increase as the predicted birthweight increases.



## Model comparison

I will compare my model with two others:  

a) including length at birth and gestational age as predictors
b) including head circumference, length, sex, and all interactions

```{r blength_gestage_model}
blength_gaweeks_fit <-
  lm(bwt ~ blength + gaweeks, data = birthweight_df) 

blength_gaweeks_fit %>% broom::tidy()
blength_gaweeks_fit %>% broom::glance()

```


```{r head_length_sex_interaction_model}
head_length_sex_fit <-
  lm(bwt ~ bhead + blength + babysex + (bhead * blength) + (blength * babysex) + (bhead * babysex) +
     (bhead * blength * babysex), data = birthweight_df) 

head_length_sex_fit %>% broom::tidy()
head_length_sex_fit %>% broom::glance()
```

## Cross-validated prediction error comparison

I will compare models in terms of the cross-validated prediction error.

I am using `crossv_mc`, which preforms training and testing splits multiple times and stores datasets in list columns.

```{r cross_validation}

bwt_cv_df <-
  crossv_mc(birthweight_df, 100) 

bwt_cv_df
```

I will fit the models and assess prediction accuracy by obtaining RMSEs using `purr::map` and `purr::map2`.

```{r model_rmse}

bwt_cv_df <- 
  bwt_cv_df %>% 
  mutate(smoken_menarche_fit    = map(train, ~lm(bwt ~ smoken + menarche, data = .x)),
         blength_gaweeks_fit    = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         head_length_sex_fit    = map(train, ~lm(bwt ~ bhead + blength + babysex + (bhead * blength) + (blength * babysex) + (bhead * babysex) +
     (bhead * blength * babysex), data = .x))) %>% 
  mutate(rmse_smoken_menarche   = map2_dbl(smoken_menarche_fit, test, ~rmse(model = .x, data = .y)),
         rmse_blength_gaweeks   = map2_dbl(blength_gaweeks_fit, test, ~rmse(model = .x, data = .y)),
         rmse_head_length_sex   = map2_dbl(head_length_sex_fit, test, ~rmse(model = .x, data = .y)))

```

The plot below shows the distribution of RMSE values for my proposed model and the two comparison models.

```{r rmse_plot}
bwt_cv_df %>% 
  dplyr::select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
    geom_violin() +
    labs(
      title = "Distribution of RMSE values",
      x = "Regression model",
      y = "RMSE"
    ) 
```

This plot compares prediction error distributions across methods.      
 
The head circumference, length, and sex model with all interactions provides the best fit. My smoking and maternal menarche age model has the largest prediction errors, although it may serve to address a more actionable or meaningful hypothesis.  